### 2022-10-11 arxiv/paper.tex
arxiv version

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{New and simplified manual controls for projection and slice
tours, with application to exploring classification boundaries in high
dimensions}

\author{
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage{setspace}\doublespacing
\usepackage{graphicx}
\usepackage{nicematrix}

\begin{document}
\maketitle

\begin{abstract}
This paper describes new user controls for examining high-dimensional
data using low-dimensional linear projections and slices. A user can
interactively change the contribution of a given variable to a
low-dimensional projection, which is useful for exploring the
sensitivity of structure to particular variables. The user can also
interactively shift the center of a slice, for example, to explore how
structure changes in local subspaces. The Mathematica package as well as
example notebooks are provided, which contain functions enabling the
user to experiment with these new manual controls, with one specifically
for exploring regions and boundaries produced by classification models.
The advantage of Mathematica is its linear algebra capabilities, and
interactive cursor location controls. Some limited implementation has
also been made available in the R package \tt{tourr}. 
\end{abstract}

\keywords{
    data visualisation; grand tour; statistical computing; statistical
    graphics; multivariate data; dynamic graphics
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

From a statistical perspective 3D is a rare data dimension, so unlike in
most 3D rotation computer graphics applications, the more useful methods
for data analysis need to work for arbitrary dimension. A good approach
is to show projections from an arbitrary dimensional space to create
dynamic data visualizations called \emph{tours}. Tours involve views of
high-dimensional (\(p\)) data with low-dimensional (\(d\)) projections.
In his original paper on the grand tour, Asimov (1985) provided several
algorithms for tour paths that could theoretically show the viewer the
data \emph{from all sides}. Prior to Asimov's work, there were numerous
preparatory developments including Fisherkeller, Friedman, and Tukey
(1974)'s PRIM-9. PRIM-9 had user-controlled rotations on coordinate
axes, allowing one to manually tour through low-dimensional projections.
(A video illustrating the capabilities is available through video
library of ASA Statistical Graphics Section (2022).) Steering through
all possible projections is impossible, unlike Asimov's tours which
allows one to quickly see many, many different projections. After Asimov
there have been many tour developments, which are summarized in Lee et
al. (2021).

One such direction of work develops the ideas from PRIM-9, to provide
manual control of a tour. Cook and Buja (1997) describe controls for 1D
(or 2D) projections, respectively in a 2D (or 3D) manipulation space,
allowing the user to select any variable axis, and rotate it into, or
out of, or around the projection through horizontal, vertical, oblique,
radial or angular changes in value. Spyrison and Cook (2020) refined
this algorithm and implemented it to generate radial tour animation
sequences.

Manual controls are especially useful for assessing sensitivity of
structure to particular elements of the projection. There are many
places where it is useful. In exploratory data analysis, where one sees
clusters in a projection, one may ask whether some variables can be
removed from the projection without affecting the clustering. For
interpreting models, one can reduce or increase a variable's
contribution to examine the variable importance. Having the user
interact with a projection is extremely valuable for understanding
high-dimensional data. However, these algorithms have two problems: (1)
the pre-processing of creating a manipulation space overly complicates
the algorithm, (2) extending to higher dimensional control is difficult.

Another potentially useful manual control, is to allow the user to
choose the position of the center of a slice. The slice tour was
introduced in Laa, Cook, and Valencia (2020). It operates by converting
the projection plane into a slice, by removing or de-emphasizing points
that are further than a fixed orthogonal distance from the plane. The
projection plane is usually thought of as passing through the center of
the data. Manual control would allow the user to change the position of
the center point, by shifting it along a coordinate axis, while keeping
the orientation of the projection plane fixed. The purpose would be to
explore how or if the shape of the data, in the space orthogonal to the
projection, changes as one gets away from the center. It would also
allow the user to interactively decide on the thickness of the slice.

This paper explains the new manual controls for projection and slice
tours. The next section describes the new algorithm for manual control,
for both projections and slices. The use of these methods is illustrated
to compare and contrast boundaries constructed by different classifiers.
The software section describes a mathematica package that is used for
the application, and describes the interactive environment that would be
desirable within R as new technology becomes available. The paper is
accompanied by an appendix with more details and adjustments to the
manual controls, and three Mathematica notebooks that can be used to
reproduce the application.

\hypertarget{sec:method}{%
\section{How to construct a manual tour}\label{sec:method}}

A manual tour allows the user to alter the coefficients of one (or more)
variables contributing to a \(d\) dimensional projection. The initial
ingredients are an orthonormal basis (\(A_{p\times d}\)) defining the
projection of the data, and a variable id (\(m \in \{1, ..., p\}\))
specifying which coefficient will be changed. A method to update the
values of the component (\(m^{th}\) row of \(A_{p\times d}\)) of the
controlled variable \(V_m\) is then needed.

\hypertarget{existing-methods}{%
\subsection{Existing methods}\label{existing-methods}}

The methods for updating component values in Cook and Buja (1997) (and
utilized in Spyrison and Cook (2020)) are prescribed primarily for a 2D
projection, to take advantage of (then) newly developed 3D trackball
controls made available for computer gaming. The first step was to
construct a 3D manipulation space from a 2D projection. In this space,
the coefficient of the controlled variable ranges between -1 and 1.
Movements of a cursor are recorded and converted into changes in the
values of \(V_m\) thus changing the displayed 2D projection. The
algorithm also provided constraints to horizontal, vertical, radial or
angular motions only. The construction of the manipulation space overly
complicates the manual controls, especially when considering possible
techniques that will apply to arbitrary \(d\).

\hypertarget{a-new-simpler-and-broadly-applicable-approach}{%
\subsection{A new simpler and broadly applicable
approach}\label{a-new-simpler-and-broadly-applicable-approach}}

The new approach emerged from experiments on the tour using the linear
algebra capabilities, and relatively new interactive graphics interface,
available in Mathematica (Wolfram Research, Inc. 2022). The components
corresponding to \(V_m\) are directly controlled by cursor movement,
which updates row \(m\) of \(A\). The updated matrix is then
orthonormalised.

\hypertarget{algorithm}{%
\subsubsection{Algorithm}\label{algorithm}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Provide \(A\), and \(m\). (Note that \(m\) could also be automatically
  chosen as the component that is closest to the cursor position.)
\item
  Change values in row \(m\), for example, if \(d=2\) gives \[
  A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
                             \vdots & \vdots \\
                             a^*_{m1} & a^*_{m2}\\
                             \vdots & \vdots \\
                             a_{p1} & a_{p2} 
       \end{array}\right].
  \] \noindent A large change in these values would correspond to making
  a large jump from the current projection. Small changes would
  correspond to tracking a cursor, making small jumps from the current
  projection.
\item
  Orthonormalise \(A^*\), using Gram-Schmidt.

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    Normalise \(\boldsymbol{a}^*_1\) and \(\boldsymbol{a}^*_2\).
  \item
    \(\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1\).
  \end{enumerate}
\end{enumerate}

This algorithm will produce the changes to a projection as illustrated
in Figure \ref{fig:manualsequence}. The controlled variable, \(V_m\),
corresponds to the black line, and sequential changes to row \(m\) of
\(A\) can be seen to roughly follow a specified position (orange dot).
Changes in the other components happen as a result of the
orthonormalisation, but are uncontrolled.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_files/figure-latex/manualsequence-1} \caption{Sequence of projections where contribution of one variable is controlled (black) is changed using unconstrained orthonormalisation. The dot (orange) indicates the chosen values for the controlled variable. It can be seen that the actual axis does not precisely match the chosen position, but it is close.}\label{fig:manualsequence}
\end{figure}

\hypertarget{refinements-to-enforce-exact-position}{%
\subsection{Refinements to enforce exact
position}\label{refinements-to-enforce-exact-position}}

The problem with the new simple method is that it is not faithful to the
precise values for \(V_m\) because the orthonormalisation will change
them. Even though these changes are for the most part imperceptible, one
may wish to avoid them and there are numerous ways that this can be
enforced, a few are detailed in the Appendix. These primarily differ in
how the remaining variables are adjusted during orthonormalisation.

\hypertarget{manual-control-for-slices}{%
\subsection{Manual control for slices}\label{manual-control-for-slices}}

To better explore the space we combine the manual controls for the
projection with manual controls for slicing. A slice is a section of the
data that is defined by a projection, a center point that is anchoring
it in the high-dimensional space and the slice thickness \(h\) (Laa,
Cook, and Valencia 2020). A data point is inside the slice if its
orthogonal distance from the projection plane (passing through the
center point) is below the thickness \(h\). This orthogonal distance is
computed in terms of the component that is normal on the projection
plane. For \(\mathbf{x}_i\) a \(p\) dimensional data point and
\(\mathbf{c}\) the center point (in the same \(p\) dimensional space) we
compute the orthogonal distance as \begin{equation}
v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}' ,
\label{eq:slice}
\end{equation} with
\(\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2\),
\(\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2\)
and \(\mathbf{a}_k, k=1,2 (=d)\) denoting the columns of the projection
matrix, \(\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)\).

\hypertarget{shifting-the-center}{%
\subsubsection{Shifting the center}\label{shifting-the-center}}

A natural starting point is to place \(\mathbf{c}\) in the center of the
data distribution, but shifting it away from the mean can provide
additional insights. In the case of a single orthogonal direction on the
projection plane we can pick a sequence of center points \(\mathbf{c}\)
in steps along that direction to move the slice and fully cover the data
space. This no longer works in higher-dimensional spaces, and we can
think of picking one direction and shifting the slice along the
component orthogonal to the projection plane.

\hypertarget{changing-the-thickness}{%
\subsubsection{Changing the thickness}\label{changing-the-thickness}}

In addition it is also useful to interactively change the slice
thickness \(h\) (also called the slice radius), in particular to find
the preferred value for exploring the input data. For guidance the
estimates of the number of points inside the slice as a function of the
original sample size \(N\) and the number of dimensions \(p\) from Laa
et al. (2022) can be used: in case of a uniform distribution inside a
sphere of radius \(R\) a slice with thickness \(h\) will contain \(N_S\)
points, with \begin{equation}
N_S(h, p, R, N) = \frac{N}{2} \left(\frac{h}{R}\right)^{p-2} \left(p - (p-2)\left(\frac{h}{R}\right)^{2}\right).
\label{eq:count}
\end{equation}

\hypertarget{sec:implementation}{%
\section{Software}\label{sec:implementation}}

The implementation of the manual tour as suggested here requires the
visualization of the current projection in terms of an axis display (see
Figure \ref{fig:manualsequence} for an example). This display should be
able to track the mouse position and adjust the projection based on the
user selection. The user can select one axis (corresponding to one
variable) by clicking on it, and then adjust the position of that axis
by dragging it during the click event. In practice the closest axis will
be selected, and the dragging results in tracking of the mouse position
which will update the projection in small steps, such that no further
interpolation is required.

The interactions in the axis display need to be mapped back onto the
current projection matrix, which will then be orthonormalized before
feeding back into the axis display. A second, linked display shows the
projected data in sync with the updates from the axis display.

In addition, we also want to be able to look at slices of the data and
select slicing parameters interactively. Here we have implemented a
switch to change between projection and slice view, a slider that
adjusts the slice thickness \(h\) (dubbed \texttt{height} in the
\texttt{Mathematica} package) and a numeric input to specify the center
point \(c\) explicitly. The display will update based on these inputs,
in particular the view will jump to a new slice when \(c\) is changed
(though an interpolation might also be useful).

One might implement such an interactive interface via R Shiny, but in
particular the tracking of the mouse position in small increments might
pose a challenge. Here we found the dynamic graphics interface from
Mathematica to be especially useful. In the following we first describe
the relevant Mathematica functionalities and how they were used in the
implementation, followed by a brief sketch of how the new approaches are
implemented in the \texttt{tourr} package. There is hope that better
interactivity will be available for R soon, which might allow for
implementation of the techniques developed in Mathematica.

\hypertarget{mathematica-package}{%
\subsection{Mathematica package}\label{mathematica-package}}

Mathematica provides much utility and versatility, such as inbuilt data
visualization, data manipulation and analysis, dynamic functionality,
and symbolic and numeric computation. Most of this inbuilt functionality
is user-friendly and described in the Mathematica documentation;
typically, numerous examples are provided within the documentation, and
some possible issues are outlined there.

Importantly for our work, it is relatively simple to create dynamic
objects via the inbuilt commands \texttt{Manipulate} or
\texttt{DynamicModule}. Control objects, such as sliders, locator panes,
and input fields, can then be used on dynamic variables, and when there
is a change in the dynamic variable the dynamic objects which contain
that variable will be updated. This is the essential ingredient for this
implementation of the manual tour.

The most relevant inbuilt Mathematica function for our purpose is the
\texttt{LocatorPane}. This creates a region on the screen where the
position of the mouse is captured and then converted to input that
updates the graphics functions, enabling the manual navigation of the
tour.

The examples presented in the application section below (and attached in
notebook format in the supplementary material) all use our primary new
function, \texttt{SliceDynamic}. This function typically accepts grouped
data in the form of a matrix where the second last column details the
name of the group and the last column details the group index (which can
be one if there is only one group). After specifying the initial slice
thickness and the slice range, the user is presented with an interactive
display in which the control objects appear on the left and the slice
visualization appears on the right. The user can change the orientation
of the slice via the locator pane, which changes the projection matrix;
a slider controls the slice thickness; and there is an input field,
which changes the slice center. The user can also change the appearance
of the plot by zooming into the center or changing the point size with
the sliders provided. It is worth noting that this zoom works best with
data that has been centered and scaled. The projected data can be
displayed to contrast it with the slice by ticking a box. The explicit
projection matrix can also be displayed via another checkbox. This is
especially useful when we wish to import a projection that was
identified as interesting in the manual exploration into a later study,
as shown in the applications below.

Details about the implementation and usage instructions are given in the
Appendix.

\hypertarget{extensions-to-the-r-package-tourr}{%
\subsection{\texorpdfstring{Extensions to the R package
\texttt{tourr}}{Extensions to the R package tourr}}\label{extensions-to-the-r-package-tourr}}

The R package tourr (Wickham et al. 2011) provides numerous types of
tours. An additional function, \texttt{radial\_tour()} has been
constructed that will generate a sequence of projections that will
decrease the coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to
the original values. It is applicable for any projection dimension,
\(d\). Changing slice center manually can be accomplished with the new
function, \texttt{manual\_slice()}. This changes the value of the center
point (\(\mathbf{c}\)) of the slice, along a selected variable axis. It
slices from the center of the data out to one side, back to the center,
out to the opposite side, and then back to the center. The appendix
contains a discussion of a possible visual guide for the slice center
location.

\hypertarget{sec:examples}{%
\section{Application}\label{sec:examples}}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{paper_files/figure-latex/penguins-scatmat-1}

}

\caption{Scatterplot matrix of the (standardised) penguins data. The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass.}\label{fig:penguins-scatmat}
\end{figure}

To illustrate the usefulness of the manual controls we use the 4D
penguins data (Horst, Hill, and Gorman 2020) and look at classification
models following Wickham, Cook, and Hofmann (2015). We will show how
classification boundaries can be explored and better understood on
projections and slices through 4D space. Figure
\ref{fig:penguins-scatmat} shows a scatterplot matrix of this data.
There are four variables
(\texttt{bl\ =\ bill\_length\_mm,\ bd\ =\ bill\_depth\_mm,\ fl\ =\ flipper\_length\_mm,\ bm\ =\ body\_mass\_g})
measuring the size of the penguins from three species (Adelie, Chinstrap
and Gentoo). The scatterplot matrix shows that the three species appear
to be likely separable, and that at least the Gentoo can be
distinguished from the other two species when \texttt{bd} is paired with
\texttt{fl} or \texttt{bm}. The steps for exploring boundaries in this
example are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Build your classification model.
\item
  Predict the class for a dense grid of values covering the data space.
\item
  Examine projections, using a manual tour so that the contribution of
  any variable is controlled.
\item
  Slice through the center, to explore where the boundaries will likely
  meet.
\item
  Move the slice by changing the center in the direction of a single
  variable to explore the extent of a boundary for a single group
  relative to a variable.
\end{enumerate}

\hypertarget{constructing-the-4d-prediction-regions}{%
\subsection{Constructing the 4D prediction
regions}\label{constructing-the-4d-prediction-regions}}

We use the \texttt{classifly} package (Wickham 2022) to generate
predictions across the 4D cube spanned by the data, with two
classification models: linear discriminant analysis (LDA) and random
forest (RF). Both the data and the model points in the grid are centered
and scaled (standard deviation \(= 1\)).

\hypertarget{exploring-projections-manually}{%
\subsection{Exploring projections
manually}\label{exploring-projections-manually}}

We start by exploring the projections of the model prediction. Figure
\ref{proj1} summarizes the process. Rotating manually the view, we can
visualize the location of predictions for each of the three species, and
also get a sense of the difference between the two models. To illustrate
this difference, we have manually rotated the projection for the RF
model (left plot) to identify a view that shows the non-linear but
block-type structure that is typical of this type of model. This
particular projection (\(A_1\)) is exported so that it can later be used
to show the LDA model (middle plot) and the actual data (right plot).
What can be seen is the linearity of the LDA model, where the boundaries
are linear and oblique to the variable axes. And, interestingly, this
particular projection of the original data shows very distinct clusters
of the three species. That means, the obscuring of the boundaries
between groups for both of the models is driven by what is happening in
the orthogonal space to the plane of the selected projection.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left). We can clearly see a block structure, while the LDA model (middle) produces linear boundaries. The three groups are nicely separated in this projection of the data (right).}
\label{proj1}
\end{figure*}

This example can be reproduced with the code in
penguins\_exploring\_manually.nb and a run through is shown in the video
at \url{https://vimeo.com/747585410}.

\hypertarget{slicing-through-the-center}{%
\subsection{Slicing through the
center}\label{slicing-through-the-center}}

We now continue the investigation by slicing orthogonally to the
projection \(A_1\). For both models we look at a thin (\(h=0.5\)) slice
through the center, \(\mathbf{c}^0 = (0,0,0,0)\). At first we explore
how changing the projection away from \(A_1\) can help with
understanding the boundary better. For our example, notice that \(A_1\)
does not contain any contribution from the second variable
(\texttt{bd}), so we will first rotate this variable into the view.

Figure \ref{slice1} shows snapshots of the exploration. The top row is
the initial projection (\(A_1\) and \(\mathbf{c}^0\)), and the bottom
row is a later projection containing more of \texttt{bd} (new projection
\(A_2\) while keeping the center fixed at \(\mathbf{c}^0\)). The columns
correspond to the two models, with RF on the left and LDA on the right.

Both models have some overlap at the center for the thin slice based on
\(A_1\) (despite the small value for \(h\)). The second slice (based on
\(A_2\)) mostly resolves this overlap and reveals the primary
differences between the models. The boundary between Adelie (blue) and
Chinstrap (yellow) is very similar for both models. The boundary between
Chinstrap and Gentoo (green) is where they differ.

The RF is almost straight in this view, as something we might expect
from a tree model where splitting occurs on single variables only.
However, it is straight in a combination of the second and third
variable (\texttt{bd} and \texttt{fl}). By examining the scatterplot
matrix, we can understand how this boundary was built: on each of
\texttt{bd} and \texttt{fl} it is possible to cut on a value where most
of the Gentoo penguins are different from the other two species for any
tree, and likely only one is needed. Thus the forest construction is
providing a sample of trees that use one variable or the other variable
to split, producing the blocky boundary on the combination of the two.

In contrast, the LDA boundary has an oblique split between the two, and
roughly divides the space into three similarly sized areas. It is almost
like a textbook illustration of how LDA works for two variables (2D)
where by assuming the clusters have equal variance-covariance, it places
a boundary half-way between the group means.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.45\textwidth]{figures/slice1_rf.png}
\includegraphics[width=0.45\textwidth]{figures/slice1_lda.png}}
\centerline{\includegraphics[width=0.45\textwidth]{figures/slice2_rf.png}
\includegraphics[width=0.45\textwidth]{figures/slice2_lda.png}}
\caption{Comparing slices based on two projections $A_1$ (top row) and $A_2$ (bottom row), for the two models RF (left) and LDA (right). With $A_1$ we see two groups overlap (green - Gentoo with yellow - Chinstrap), while the rotation to $A_2$ results in clear boundaries inside the slice. The boundary between Adelie (blue) and Chinstrap (yellow) is similar for both models but very different between Chinstrap and Gentoo (green).}
\label{slice1}
\end{figure*}

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.45\textwidth]{figures/proj2_data.png}}
\caption{Projection of the data based on $A_2$. Compared to projecting onto $A_1$ we see that the green observations (Gentoo) are more separated from the other two species.}
\label{proj2}
\end{figure*}

Finally, to determine which model describes the boundaries better,
compare them with the \(A_2\) projection of the data (Figure
\ref{proj2}). Gentoo is more separated from the other two in this
projection, and one can imagine that the trees in the forest has
greedily grasped any one of many places to make a split to separate the
group. It might be argued though the that RF boundary is cut too close
to the Chinstrap species, and might lead to some unnecessary
misclassification with new data. The LDA boundary is better placed for
all species.

This example can be reproduced with the code in
penguins\_slicing\_through\_center.nb and a run through is shown in the
first part of the video at \url{https://vimeo.com/747590472}.

\hypertarget{shifting-the-slice-center}{%
\subsection{Shifting the slice center}\label{shifting-the-slice-center}}

We have seen that starting from \(A_1\) using the manual controls to
change the contribution of the second variable we could find a clear
separation boundary indicating the relation between this variable (bill
depth) and the Gentoo penguin species. Instead of rotating to a
different projection, we might also change the view by moving the slice
along one axis in the 4D space. Here we will continue our exploration of
the dependence on \texttt{bd} and move the slice defined by \(A_1\) to
either large positive or negative values (\(c^{\pm}_{bd} = \pm 1.5\)
after centering and scaling). Thus we define
\(\mathbf{c}^{\pm} = (0, c^{\pm}_{bd}, 0, 0)\). Here we will also look
at slices of the observed data points, using a thicker slice (\(h=1.5\))
to capture enough points in a given view.

We start by a comparison of the two models and the data distribution
shifting to \(\mathbf{c}^{+}\), thus the slice is localized towards high
values of bill depth in Figure \ref{slice1p}. We can see that all three
slices (the two models and the data) contain almost no points from the
third class (green, Gentoo), and that the decision boundary between the
two models is very similar.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_p_rf.png}
\includegraphics[width=0.32\textwidth]{figures/slice1_p_lda.png}
\includegraphics[width=0.32\textwidth]{figures/slice1_p_data.png}}
\caption{Shifting the slice center in the positive direction of bill depth produces regions that have no Gentoo (green). The two models have similar boundaries except that the RF is more of a step function.}
\label{slice1p}
\end{figure*}

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.32\textwidth]{figures/slice1_m_rf.png}
\includegraphics[width=0.32\textwidth]{figures/slice1_m_lda.png}
\includegraphics[width=0.32\textwidth]{figures/slice1_m_data.png}}
\caption{Shifting the slice center in the negative direction of bill depth produces regions that are mostly Gentoo (green). The two models have very different boundaries: the nonlinear potential of RF can be seen here where still some subspaces would predict to be Adelie and Chinstrap.}
\label{slice1m}
\end{figure*}

A more interesting comparison is found for \(\mathbf{c}^{-}\), thus the
slice localized towards low values of bill depth, shown in Figure
\ref{slice1m}. The RF model (left) predicts all three species within
this slice, with an interesting boundary for the third class (green,
Gentoo). On the other hand the LDA model (middle) predominantly predicts
the third class within the slice, this appears to be enforced through
the linear structure of the model. Looking finally at the thick slice
through the data we see that there are primarily observations from this
class within the slice we can conclude that the two models have filled
in the ``empty'' space (where we do not have any training observations)
in very different ways and according to what we might expect given the
model structure.

Finally it is interesting to compare the slice views to the projection
of the models seen in Fig. \ref{proj1} to better understand how the
boundaries change along the \texttt{bd} direction and where the
differences in the projections come from.

This example can be reproduced with the code in
penguins\_shifting\_slice\_center.nb and a run through is shown in the
second part of the video at \url{https://vimeo.com/747590472}.

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

This short note has described new technology for manually interacting
with tours. The manual control of projection coefficients is most useful
for assessing variable importance to perceived structure, but can be
generally used for steering a viewer through high-dimensional space.
Changing the center of a slice manually enables exploring the space
orthogonal to a projection, specifically in the direction of a single
variable. This is a new tool, and as shown in the example, when used
with the slice tour can be very useful for understanding boundaries of
classifiers. This is also likely a useful method for examining
low-dimensional non-linear manifolds, and also functions of multiple
parameters.

Mathematica provided a useful sandbox to experiment with the ideas
presented in this paper. Most of the data, though, is first constructed
using R, especially the classification boundaries. If new technology for
interactive graphics becomes available in R it would be useful to create
a tighter coupling of models and visualization to allow exploring and
comparing fits.

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

The authors gratefully acknowledge the support of the Australian
Research Council and the ResearchFirst undergraduate research program at
Monash University. The paper was written in \texttt{rmarkdown} (Xie,
Allaire, and Grolemund 2018) using \texttt{knitr} (Xie 2015). The
scatterplot matrix of the penguins data was produced by the GGally
(Emerson et al. 2013) package built on ggplot2 (Wickham 2016) graphics.
We thank the Institute of Statistics, BOKU, for their hospitality while
part of this work was conducted.

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

The source material and animated gifs for this paper are available at
\url{https://github.com/uschiLaa/mmtour}.

The supplementary materials include:

\begin{itemize}
\tightlist
\item
  The Mathematica source code defining the new functions in mmtour.wl.
\item
  Three Mathematica notebooks with the code used in the application
  (corresponding to the three subsections).
\item
  Appendix with additional details about manual controls, and
  Mathematica functions.
\item
  Animations illustrating the manual tour and slicing, matching the
  static figures in the paper. These are also available at
  \url{https://vimeo.com/747585410} and
  \url{https://vimeo.com/747590472}.
\end{itemize}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ASA22}{}}%
ASA Statistical Graphics Section. 2022. {``Video {L}ibrary.''}
\url{https://community.amstat.org/jointscsg-section/media/videos}.

\leavevmode\vadjust pre{\hypertarget{ref-As85}{}}%
Asimov, D. 1985. {``{T}he {G}rand {T}our: {A} {T}ool for {V}iewing
{M}ultidimensional {D}ata.''} \emph{SIAM Journal of Scientific and
Statistical Computing} 6 (1): 128--43.
https://doi.org/\url{https://doi.org/10.1137/0906011}.

\leavevmode\vadjust pre{\hypertarget{ref-cook_manual_1997}{}}%
Cook, Dianne, and Andreas Buja. 1997. {``Manual {Controls} for
{High}-{Dimensional} {Data} {Projections}.''} \emph{Journal of
Computational and Graphical Statistics} 6 (4): 464--80.
\url{https://doi.org/10.2307/1390747}.

\leavevmode\vadjust pre{\hypertarget{ref-GGally}{}}%
Emerson, John W., Walton A. Green, Barret Schloerke, Jason Crowley,
Dianne Cook, Heike Hofmann, and Hadley Wickham. 2013. {``The Generalized
Pairs Plot.''} \emph{Journal of Computational and Graphical Statistics}
22 (1): 79--91. \url{http://www.jstor.org/stable/43304816}.

\leavevmode\vadjust pre{\hypertarget{ref-tukey}{}}%
Fisherkeller, M. A., J. H. Friedman, and J. W. Tukey. 1974. {``{PRIM-9},
an Interactive Multidimensional Data Display and Analysis System.''} In
\emph{The Collected Works of John w. Tukey: Graphics 1965-1985, Volume
v}, edited by William S. Cleveland, 340--46.

\leavevmode\vadjust pre{\hypertarget{ref-penguins}{}}%
Horst, Allison Marie, Alison Presmanes Hill, and Kristen B Gorman. 2020.
\emph{Palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data}.
\url{https://allisonhorst.github.io/palmerpenguins/}.

\leavevmode\vadjust pre{\hypertarget{ref-sectionpursuit}{}}%
Laa, Ursula, Dianne Cook, Andreas Buja, and German Valencia. 2022.
{``Hole or Grain? A Section Pursuit Index for Finding Hidden Structure
in Multiple Dimensions.''} \emph{Journal of Computational and Graphical
Statistics} 0 (0): 1--14.
\url{https://doi.org/10.1080/10618600.2022.2035230}.

\leavevmode\vadjust pre{\hypertarget{ref-slicetour}{}}%
Laa, Ursula, Dianne Cook, and German Valencia. 2020. {``A Slice Tour for
Finding Hollowness in High-Dimensional Data.''} \emph{Journal of
Computational and Graphical Statistics} 29 (3): 681--87.
\url{https://doi.org/10.1080/10618600.2020.1777140}.

\leavevmode\vadjust pre{\hypertarget{ref-lee2021}{}}%
Lee, Stuart, Dianne Cook, Natalia da Silva, Ursula Laa, Earo Wang, Nick
Spyrison, and H. Sherry Zhang. 2021. {``Advanced {Review}: {The}
{State}-of-the-{Art} on {Tours} for {Dynamic} {Visualization} of
{High}-Dimensional {Data}.''} \emph{arXiv:2104.08016 {[}Cs, Stat{]}},
April. \url{http://arxiv.org/abs/2104.08016}.

\leavevmode\vadjust pre{\hypertarget{ref-spyrison_spinifex_2020}{}}%
Spyrison, Nicholas, and Dianne Cook. 2020. {``Spinifex: An {R} {Package}
for {Creating} a {Manual} {Tour} of {Low}-Dimensional {Projections} of
{Multivariate} {Data}.''} \emph{The R Journal} 12 (1): 243.
\url{https://doi.org/10.32614/RJ-2020-027}.

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\leavevmode\vadjust pre{\hypertarget{ref-classifly}{}}%
---------. 2022. \emph{Classifly: Explore Classification Models in High
Dimensions}. \url{https://CRAN.R-project.org/package=classifly}.

\leavevmode\vadjust pre{\hypertarget{ref-sam11271}{}}%
Wickham, Hadley, Dianne Cook, and Heike Hofmann. 2015. {``Visualizing
Statistical Models: Removing the Blindfold.''} \emph{Statistical
Analysis and Data Mining: The ASA Data Science Journal} 8 (4): 203--25.
\url{https://doi.org/10.1002/sam.11271}.

\leavevmode\vadjust pre{\hypertarget{ref-tourr}{}}%
Wickham, Hadley, Dianne Cook, Heike Hofmann, and Andreas Buja. 2011.
{``{tourr}: An {R} Package for Exploring Multivariate Data with
Projections.''} \emph{Journal of Statistical Software} 40 (2): 1--18.
https://doi.org/\url{http://dx.doi.org/10.18637/jss.v040.i02}.

\leavevmode\vadjust pre{\hypertarget{ref-Mathematica}{}}%
Wolfram Research, Inc. 2022. \emph{Mathematica, {V}ersion 13.1}.
Champaign, IL. \url{https://www.wolfram.com/mathematica}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed.
Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
Xie, Yihui, Joseph J. Allaire, and Garrett Grolemund. 2018. \emph{R
Markdown: The Definitive Guide}. Chapman; Hall/CRC.
\url{https://bookdown.org/yihui/rmarkdown}.

\end{CSLReferences}

\bibliographystyle{unsrt}
\bibliography{biblio.bib}

\end{document}